Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data

### 2018-10-08 paper.tex

        \section{Tool interfaces}\label{mod}
            \tristan{Since this section is about interfaces, I would
            rather call it ``Interfaces" or ``Tool interfaces".}
            Often times, users will want to integrate existing tools into their 
            workflows. In order for workflows to communicate effectively with
            the tools, it is required that an interface be created. Workflow 
            interfaces may range from simple command-line calls, to more 
            complex ones that track inputs, outputs and other parameters. 
            Furthermore, they ensure
            that all necessary parameters are correctly defined and in the 
            correct order,  capture all of the 
            standard input/output and create folders to encapsulate task inputs
            and outputs. Workflow engines typically rely on more complex types 
            of interfaces in order to ensure correct functioning of the pipeline
            . As these interfaces can be rather complex, it is necessary for the 
            workflow engine to provide an intuitive way for users to create 
            interfaces for their custom tools. Workflow interface can be 
            divided into two groups: those that are described using a 
            programming language and those that are described using DSL.

            Workflow engines that utilize programming languages include: SPM,
            One of the earliest neuroimaging workflow engines, 
            Early neuroimaging workflow engines, such as SPM, are monolithic, 
            and therefore, do not provide a default interface by which users 
            can integrate their tools\tristan{Users can still integrate tools as Matlab functions
            doing system calls}. Workflow engines, such as AA, were 
            created to enable engines such as SPM to be more modular. AA 
            allows users to define their tasks through XML specification. AA's 
            XML tasklist is simply an ordered list of tasks which also includes
            details on analysis-specific parameters. Dependencies are
            inferred by the program through analysis of the module interfaces.
            Module interfaces consist of MATLAB user script and an XML 
            descriptor. The descriptor contains details on the input and output
            data stream types (e.g. epi, structural, dicom\_header) 
            and the desired domain (e.g. study, subject, session). During 
            execution, each task is given its own unique folder to operate on. 
            The engine will copy data to and extract data from this folder.

            LONI, FastR, Taverna, Pegasus and Galaxy are other examples of 
            workflow engines that 
            use an XML interface to describe CLIs. LONI enables generation of 
            the XML-based module interface through its GUI. The module interface
            consists of details on \note{from website} module metadata (e.g.
            authors, package, version, name) and citation information, 
            command-line syntax, parameter and file types, dependencies and 
            output and error handling.

            Tool addition in Taverna is primarily performed through the user 
            interface, but can also achieved through the creation of an XML
            descriptor. Taverna differs from other workflow engines in that it
            execute generic-style services (i.e. WSDL and REST) in addition to
            other types of Web Services, as well a external CLIs, R scripts and
            Java services. To interface with a CLI, the XML file must contain 
            the command to execute, file inputs and outputs (if required), if 
            standard input, output and error streams must be conserved and the 
            location where the command should execute\note{from website}.

            Pegasus is another workflow engine that uses an XML format to 
            define the CLI tools interface. The XML interfaces are typically 
            incorporated directly into the workflow DAX definition. However, a 
            separate multiline text-based 'Transformation Catalogue' file can
            be provided. Information encoded within the interfaces include
            tool identifier (typically namespace::version), the URL or filepath
            to the executable, site identifier of the executable's host, 
            executable type (e.g. installed or staged) and the profile. Pegasus
            can also execute containerized tools \tristan{how about the other ones?}. For this, a container 
            identifier, image URL, site identifier of the image, and container
            profiles must also be defined \note{from website}.

            While Galaxy is primarily a GUI-based application, an XML tool
            configuration file is required to add the tool to the application.
            The tool wrappers schema is very rich allowing for many tool 
            features to be defined. Similarly to many other workflow engine tool
            wrapper, the configuration file contains details on the tool name, 
            type, id, versions, command and input and output specifications. It
            also enables the execution of containerized tools. Furthermore,
            Galaxy enables users to specific actions to be performed on the 
            output files once created and define test cases.

            While many worfklow engines use XML or JSON to describe their tools,
            others wrap their tools in an interface defined by a
            programming/scripting language. Nipype, Pydpiper and Toil all use
            Python to create their wrappers; PSOM wrappers are written in
            MATLAB/Octave; OpenMOLE uses Java/Scala; and Make interfaces are
            written in Shell Script.

            Nipype's wrapper, written in Python, can encapsulate CLIs, SPM, 
            MATLAB and Python scripts. Attributes and methods that can be 
            defined for all interfaces include InputSpec, OutputSpec and a 
            Python dictionary of all expected outputs. InputSpec and OutputSpec
            are two Nipype classes. The former defines the input parameters of 
            a command and the second defines the outputs that are possibly 
            generated by the command. For CLIs, it is necessary to define the
            command. Additionally, Nipype will start X virtual framebuffer 
            before executing a CLI, if specified \tristan{you need to explain why}. For SPM interfaces, the job 
            type and name must be specified. Interfaces to MATLAB require 
            definition of the run function, while Python interfaces require
            the definition of the \textit{\_run\_interface} function. \note{from
            website}.

            Pydpiper's tools can either be custom Python code or CLIs. The 
            \textit{PipelineStage} and \textit{CmdStage} are used to 
            encapsulate the tools. The \textit{CmdStage} class is specifically 
            used to interface with CLIs, and thus, accepts a parsed array of the
            command and its parameters as input. Atoms, typically used to wrap
            around MINC file-format related tools, inherit from the
            \textit{CmdStage} class, and require the input MINC file (string or
            file handler) to be provided. A output file can also be specifiec, 
            but it must be of the same format as the MINC file (string or file 
            handler). Additionally, it is possible to pass multiple optional 
            arguments to the Atom.

            Toil is another workflow engine that describes their tool interfaces
            in Python. The wrappers, also known as \textit{Jobs} enable users to
            interface with custom Python code or Dockerized \note{cite Docker} 
            CLIs. Interfaces can be constructed either by creating a class that
            inherits from the \textit{Job} class or through the use of a 
            built-in function that converts the user-defined function (UDF) to a 
            Toil \textit{Job}. To convert, all that is required is the UDF,
            however, the user may also choose to include details on memory, 
            number of cores and disk space required. For Dockerized tools, Toil
            provides a built-in function that accepts the tool name, working dir
            and parameters of the tool.

            PSOM's tool interfaces are known as \textit{Bricks}. \textit{Bricks}
            can interface with any MATLAB function or CLI. Bricks require the
            definition of input and output files, in addition to a structure of
            program options. A useful feature found in PSOM is the ability 
            to perform a ``dry-run" of the Brick without executing the command.

@@

            Unlike scientific workflow engines, BigData frameworks do not 
            provide default interfaces for the execution of tools written in 
            languages different from their own. Users defining their workflows
            in Dask or the Spark and MapReduce using the Python API integrate
            Python UDFs in their workflow. While Dask users can reference their 
            UDF of Python function directly in their workflow, Spark users
            must define a lambda function within an action or transformation 
            that executes their UDF or library on the data. Similarly to Spark,
            functions referenced in MapReduce must be called from within a Map
            or Reduce task. MapReduce can additionally reference Java 
            code and Spark can call on code also written in Scala, Java and R.
            \note{verify if it's possible to mix-and-match different libraries
            ...don't think so}

            For all workflow engines that do not, by default, interface with 
            command-line tools or containers, there are libraries that can be
            easily integrated to achieve that. Boutiques, for instance, is a 
            Python library that can execute command-line or containerized tools
            defined within a JSON schema. Furthermore, the library's richness
            compete with Galaxy's XML. Python-based BigData frameworks can 
            easily call on this library from within the workflow in order to
            communicate with the tools.

@@

        In this Chapter, we will compare scientific
        workflows to BigData frameworks on the basis of performance. 
        Section~\ref{fs} will look at what filesystems are \weird{required} \tristan{supported?} by the 
        engines. This is important because data movement of any kind is costly,
        and may be worse depending on the filesystems available. \tristan{Also, mention that neuroinformatics
        pipelines manipulate files.}
        Section~\ref{sched} will provide an overview of the schedulers 
        available and their compatibility with the workflow engines. 
        Section~\ref{other} will discuss other performance improvement 
        strategies employed by the workflow engines.

@@

            Unlike MapReduce and Spark-based workflows, Dask and scientific 
            workflows are not data-centric. They focus on the efficient 
            scheduling of tasks and rely on I/O and network transfers to
            communicate the data between nodes. With BigData workflows, such
            operations could incur significant performance penalties. 
            \weird{Filesystems}. To mitigate this and remain efficient in the face of
            BigData, Dask implements lazy evaluation and in-memory processing.
            BigData platforms such as MapReduce rely heavily on I/O, but the 
            I/O performed occurs on the compute node rather than on 
            shared storaged, making it generally more efficient than scientific
            workflows. Additionally, MapReduce uses the concept of 
            locality to ensure limit data transfers across the network. Spark
            uses data locality, lazy evaluation and in-memory computing to 
            ensure its optimal performance.
            For the efficient processing of neuroimaging BigData, it will be
            necessary for scientific worflow engines to incorporate these 
            strategies.

@@

                    \note{mention caching}

                    While Spark and Dask both rely on the concept of shared 
                    memory, Spark avoids data transfers as often as possible by
                    scheduling tasks to nodes where the data is located. Dask
                    , however, will schedule its tasks to the next available 
                    node.

                \subsubsection{Data locality}
                    Data locality, with respect to BigData frameworks, is a
                    scheduling
                    strategy in which location of the data is known by the 
                    scheduler and tasks are scheduled to the location where the
                    necessary data is located. It was one of the first main 
                    strategies implemented to make BigData processing more 
                    efficient. The reasoning behind data locality is that 
                    network transfers are expensive, particularly within 
                    commodity infrastracture, and therefore transferring large 
                    volumes of data across the network for every task would be 
                    rather expensive.

@@

                    In contrast to MapReduce, Spark does not require data to 
                    be stored on a HDFS-like filesystem. It can read data from
                    any form of shared-filesystem, including Lustre. If data 
                    is extracted from a non HDFS-like filesystem, the initial 
                    loading of data will not use data locality. However, in all
                    future tasks in which the data is already loaded into an 
                    RDD, the scheduler will be able to assign task to the most
                    favorable partition dictated by delay scheduling according
                    to data locality.

@@

                    Lazy evaluation is a strategy used throughout computer 
                    science to improve performance of operations. Rather than
                    evaluating a expression immediately, it will wait until it 
                    is required in order to compute the result. This is 
                    achieved in Spark through the calling of an \textit{action}
                    on an RDD and in Dask through the calling of a 
                    \textit{compute} function on one of its data structures.
                    By using lazy evaluation, the engines are capable of 
                    constructing graphs that result in the most optimal
                    task order of execution. \tristan{And further save on data transfers}

@@

                Alluxio, formerly known as Tachyon, is an in-memory filesystem
                that stores lineage data to offset the cost of writing data to 
                disk. The assumption that Alluxio makes is that it will be less
                costly to recompute the data that to store the data to disk and
                replicate across the network for fault tolerance. However,
                unlike Spark applications, Alluxio is meant to run indefinitely
                , making it unfavourable, in some cases, to recompute the data
                from its starting point. For this, Alluxio evaluates the 
                lineage data and checkpoints the files which have been 
                repeatedly accessed (i.e. hot files). This strategy also helps 
                in avoiding the checkpointing of intermediary files.

                Triple-H leverages the heterogeneity of HPC clusters to 
                optimize the data's placement across the cluster for most 
                efficient use. It aims to reduce the I/O bottleneck of writing 
                through data buffering and caching. Triple-H has two modes of
                operation, Default and Lustre integrated,
                that allows it to execute on systems both with and without 
                access to a distributed filesystem. To further reduce I/O in
                HPC systems, Triple-H leverage's Lustre's fault tolerance 
                instead of using HDFS's replication-based fault-tolerance. 
                Triple-H improves 
                performance of cluster I/O by writing the data to
                the most optimal storage layer (RAMDisk, SSD, HDD, Lustre) 
                dictacted by the selected placement policy (i.e. Greedy or 
                Load-balanced).

                Although both of these solutions have shown to improve I/O
                performance

@@

            Scientific workflows typically rely on HPC schedulers for 
            distributed scheduling on clusters. These schedulers include SLURM,
            PBS,TORQUE, SGE, HTCondor and Mesos. Additionally, they may provide
            their own form of scheduling for computations on a single node. 
            BigData frameworks, such as Dask,
            also provide interfaces to these HPC schedulers. However, other
            frameworks such as MapReduce and Spark are limited to BigData 
            clusters such as Yet Another Resource Negotiator (YARN and MESOS).
            Spark additionally provides its own standalone scheduler. Similarly
            to scientific workflows, all three BigData frameworks provide their
            own schedulers for single-node scheduling.

@@

            Since Spark does not natively interface with HPC schedulers, 
            the use of overlay Spark clusters, similar to HoD, has been 
            investigated for Spark applications running on HPC. The reasons 
            listed for HoD's failure include poor 
            data locality when deploying a dynamic cluster on a shared HDFS 
            instance, inability to resize cluster resulting in idle nodes and 
            high cluster allocation latency. To address the issue of high 
            cluster allocation, pilot-agent scheduling strategies can be used.
            Pilot-agent scheduling strategies allow for clusters to grow 
            dynamically and to add nodes when required. This relieves users, 
            in the case of Spark, from having to wait until all nodes become 
            available to begin execution. Both papers \note{jha and baer} use 
            this strategy. Furthermore, pilot-agent scheduling could be adapted
            to enable smart cluster resizing. Conversly, maintaining data 
            locality using a shared HDFS instance will remain an issue. 
            \tristan{The section is interesting, but it lacks a bit of context: one would like to 
            understand why Spark and Hadoop developed their own scheduler instead of
            reusing the ones available on HPC. And what are the main characteristics of these
            schedulers: data locality should be mentioned here.}

@@

        \section{Evaluation of BigData Frameworks for Scientific Computing}     
            The BigData frameworks have sparked significant interest within the 
            scientific community, leading to research on the efficiency of      
            these frameworks for scientific computing. ~\cite{valduriez} found  
            that Spark scales better than expected with longer tasks. This is   
            because                                                             
            there is less I/O occurring and the scheduler is not as overloaded. 
            Additionally, ~\cite{Boubela} found that Spark in conjunction with  
            GPU acceleration greatly reduced the processing time of large       
            neuroimaging datasets.                                              
            As scientific, specifically neuroimaging workflows are typically    
            comprised of long tasks, it                                         
            is crucial that BigData frameworks scale well with tasks of longer  
            durations.

            To determine how different BigData frameworks compare in their 
            efficiency for the processing of scientific workflows, 
            ~\cite{ariel} compared Spark, Dask, SciDB, Myria and TensorFlow on
            a neuroimaging and astronomy use case. It was found that Dask, 
            Myria and Spark performed consistently better and had comparable 
            performance. Dask had poorer performance when data did not fit in
            memory. It was also found that Spark exhibited slower performance 
            when the individual images are larger due to the lack of overlap 
            between shuffle and map stages \note{revise this}. Overall it was
            found that BigData Frameworks still required some adaptations to be
            effectively used as scientific workflow engines.

    \chapter{Portability}\label{portability}
        In order for a workflow engine to be utilized, \weird{it much be simple} for 
        users to \weird{install and execute the data}. Platform-independence of both 
        the frameworks and tools, as well as, simple installation of all 
        necessary components and straightforward execution of workflows  are a 
        requirement for frameworks and their respective pipelines to be 
        portable. Without portability, users are limited to executing certain
        frameworks only or dealing with the significant overhead of making 
        these engines work on the system available to them. In this chapter, 
        we evaluate the portability of workflows on the basis of the following
        three components: 1) platform-independence, 2) tool installation and
        3) execution setup. We also examine containerization, one of the 
        strategies used to improve portability.

        \section{Platform-independence}
            Achieving platform-independence allows users to use the workflow 
            engine in any available environment. This however does not imply 
            that the tools, and as a consequence, the pipelines created will
            also be platform-independent. Many workflow engines have opted to 
            be platform independent. These workflow engines include: Nipype,
            PSOM, SPM, FASTR, LONI, AA, Pydpiper, Taverna, MapReduce, Spark and
            Dask. These workflow engines achieve platform independence through
            running on the Java Virtual Machine (JVM) or being written in 
            Python. Platform-independent engines that are Java-based include: 
            PSOM, SPM, LONI, AA, Taverna, MapReduce and Spark \tristan{this list contains Matlab
            engines}. The remainder of
            the pipelines are implemented in Python. The remainder of the 
            workflow engines (Pegasus, Make, OpenMOLE, Nextflow, Galaxy, and 
            Toil are not platform-independent, but can execute on UNIX-like 
            systems \tristan{macOS and Linux}.

            Engines built on top of a platform-independent language does not 
            necessarily imply platform independence \tristan{Grammar? Engines ... does not}. For instance, OpenMOLE is 
            built in Scala and Netflow runs on Groovy, yet neither are platform
            independent. In the case of OpenMOLE, this is due to its CARETask
            blackbox, which is only possible for Linux environments. However,
            many tools and programs available for incorporation into workflows
            are platform-dependent. For example, even though Nipype is 
            platform-independent, common tools used in Nipype modules, such as 
            FreeSurfer and FSL, are Linux-based. Moreover, HPC infrastructure 
            is usually Linux-based and it is therefore crucial that all 
            workflows are compatible with and tested on Linux systems.

>>>>

        \section{Tool interfaces}\label{mod}
            \tristan{Since this section is about interfaces, I would
            rather call it ``Interfaces" or ``Tool interfaces".}
            Often times, users will want to integrate existing tools into their 
            workflows. In order for workflows to communicate effectively with
            the tools, it is required that an interface be created. Workflow 
            interfaces may range from simple command-line calls, to more 
            complex interfaces that track inputs, outputs and other parameters. 
            Furthermore, they ensure
            that all necessary parameters are correctly defined and in the 
            correct order,  capture all of the 
            standard input/output and create folders to encapsulate task inputs
            and outputs. Early scientific workflow engines, such as SPM, are 
            monolithic. As a consequence, these workflow engines do not provide
            a complex class for interface definition and validation. Recent
            Workflow engines take a modular approach and promote sharing of data
            and pipelines. Thetypically rely on more complex types 
            of interfaces in order to ensure correct functioning of the pipeline
            . Since these interfaces can be rather complex, it is necessary for the 
            workflow engine to provide an intuitive way for users to create 
            interfaces for their custom tools. Workflow interfaces can be 
            divided into two groups: those that are described using a 
            programming language and those that are described using DSL.

            Making the choice between a programming or a markup language for 
            interface definition is not obvious. One of the main reasons for 
            selecting a programming language is that programming language is 
            more versatile than are markup language, allowing any type of 
            operation to be defined. For instance, a dynamic number of outputs
            can be generated using a programming-language based interface 
            definition. Workflow engines that utilize programming 
            languages for the 
            description of their workflow interfaces include: Nipype, PSOM
            , Pydpiper, Toil, OpenMOLE and Make interfaces.

            Nipype's wrapper, written in Python, can encapsulate CLIs, SPM, 
            MATLAB and Python scripts. InputSpec and OutputSpec are two classes 
            that can are used to describe the inputs and outputs of the 
            interface. The former defines the input parameters of 
            a command and the latter defines the outputs that are possibly 
            generated by the command. For CLIs, it is necessary to describe the
            command, such as order of parameters and required and optional parameters. 
            Apart from Input and Output Spec, a Python dictionary containing all
            expected outputs can be included.
            Before executing a CLI, Nipype can start X virtual framebuffer if 
            user interaction through a GUI is required by the tool if specified. 
            For SPM interfaces, the job 
            type and name must be specified. Interfaces to MATLAB require 
            definition of the run function, while Python interfaces require
            the definition of the \textit{\_run\_interface} function. \note{from
            website}.

            PSOM’s tool interfaces are known as \textit{Bricks}. They are a 
            little less versatile than Nipype'sinterfaces since they can only
            interface with MATLAB functions or CLIs. Bricks require the 
            definition of input and output files, in addition to a structure of 
            program options. A useful feature found in PSOM is the ability to 
            perform a “dry-run” of the Brick without executing the command.

            Pydpiper's tools can either be custom Python code or CLIs. The 
            \textit{PipelineStage} and \textit{CmdStage} are used to 
            encapsulate the tools. The \textit{CmdStage} class is specifically 
            used to interface with CLIs, and thus, accepts a parsed array of the
            command and its parameters as input. Atoms, typically used to wrap
            around MINC file-format related tools, inherit from the
            \textit{CmdStage} class, and require the input MINC file (string or
            file handler) to be provided. A output file can also be specified, 
            but it must be of the same format as the MINC file (string or file 
            handler). Additionally, it is possible to pass multiple optional 
            arguments to the Atom.

            Toil is another workflow engine that describes their tool interfaces
            in Python. The wrappers, also known as \textit{Jobs} enable users to
            interface with custom Python code or Dockerized \cite{merkel2014docker} 
            CLIs. Interfaces can be constructed either by creating a class that
            inherits from the \textit{Job} class or through the use of a 
            built-in function that converts the user-defined function (UDF) to a 
            Toil \textit{Job}. To convert, all that is required is the UDF,
            however, the user may also choose to include details on memory, 
            number of cores and disk space required. For Dockerized tools, Toil
            provides a built-in function that accepts the tool name, working dir
            and parameters of the tool.

@@

            \tristan{Users can still integrate tools as Matlab functions
            doing system calls}.
            Although DSL-based workflow interface schemas may be less versatile
            than those written using programming languages, DSL based schemas
            can be reused by any engine. This allows interface schemas to be 
            portable. AA, LONI, FastR, Taverna, Pegasus and Galaxy all 
            implement their iterface definition using a DSL.

            AA 
            allows users to define their tasks through XML specification. AA's 
            XML tasklist is simply an ordered list of tasks which also includes
            details on analysis-specific parameters. Dependencies are
            inferred by the program through analysis of the module interfaces.
            Module interfaces consist of MATLAB user script and an XML 
            descriptor. The descriptor contains details on the input and output
            data stream types (e.g. epi, structural, dicom\_header) 
            and the desired domain (e.g. study, subject, session). During 
            execution, each task is given its own unique folder to operate on. 
            The engine will copy data to and extract data from this folder.

            LONI enables generation of 
            the XML-based module interface through its GUI. The module interface
            consists of details on \note{from website} module metadata (e.g.
            authors, package, version, name) and citation information, 
            command-line syntax, parameter and file types, dependencies and 
            output and error handling.

            Tool addition in Taverna is primarily performed through the user 
            interface, but can also achieved through the creation of an XML
            descriptor. Taverna differs from other workflow engines in that it
            can 
            execute generic-style services (i.e. WSDL and REST) in addition to
            other types of Web Services, external CLIs, R scripts and
            Java services. To interface with a CLI, the XML file must contain 
            the command to execute, file inputs and outputs (if required), if 
            standard input, output and error streams must be conserved and the 
            location where the command should execute\note{from website}.

            Pegasus is another workflow engine that uses an XML format to 
            define the CLI tools interface. The XML interfaces are typically 
            incorporated directly into the workflow DAX definition. However, a 
            separate multiline text-based 'Transformation Catalogue' file can
            be provided. Information encoded within the interfaces include
            tool identifier (typically namespace::version), the URL or filepath
            to the executable, site identifier of the executable's host, 
            executable type (e.g. installed or staged) and the profile. In
            addition to UDFs and CLIs, Pegasus
            can also execute containerized tools \tristan{how about the other ones?}. For this, a container 
            identifier, image URL, site identifier of the image, and container
            profiles must also be defined \note{from website}.

            While Galaxy is primarily a GUI-based application, an XML tool
            configuration file is required to add the tool to the application.
            The tool wrappers\ schema is very rich allowing for many tool 
            features to be defined. Similarly to many other workflow engine tool
            wrappers, the configuration file contains details on the tool name, 
            type, id, versions, command and input and output specifications. It
            also enables the execution of containerized tools. Furthermore,
            Galaxy enables users to specific actions to be performed on the 
            output files once created and define test cases.

            Unlike scientific workflow engines, BigData frameworks do not 
            provide default interfaces for the execution of complex tools 
            written in 
            languages different from their own. While Dask users can reference
            UDF Python functions directly in their workflow, Spark users
            must define a lambda function within an action or transformation 
            that executes their UDF or library on the data. Similarly to Spark,
            functions referenced in MapReduce must be called from within a Map
            or Reduce task. MapReduce can additionally reference Java 
            code and Spark can call on code also written in Scala, Java and R.

            For all workflow engines that do not, by default, interface with 
            command-line tools or containers, there are libraries that can be
            easily integrated to achieve that. 
            Boutiques~\cite{doi:10.1093/gigascience/giy016}, for instance, is a 
            Python library that can execute command-line or containerized tools
            defined within a JSON schema. Furthermore, the library's richness
            compete with Galaxy's XML. Python-based BigData frameworks can 
            easily call on this library from within the workflow in order to
            communicate with the tools.

@@

        In this Chapter, we will compare scientific
        workflows to BigData frameworks on the basis of performance. 
        Section~\ref{fs} will look at what filesystems are supported \tristan{supported?} by the 
        engines. This is important because data movement of any kind is costly,
        and may be worse depending on the filesystems available. Furthermore, 
        typical neuroinformatics pipelines manipulate files, and therefore, 
        reducing the amount of I/O is necessary.\tristan{Also, mention that neuroinformatics
        pipelines manipulate files.}
        Section~\ref{sched} will provide an overview of the schedulers 
        available and their compatibility with the workflow engines. 
        Section~\ref{other} will discuss the checkpointing of data in the context
        of scientific workflows. This chapter will conclude with a discussion
        on the efficacy of Big Data frameworks for scientific computing, in
        Section~\ref{eval}.

@@

            Although MapReduce, Spark and Dask are all data-intensive worflows,
            Dask can be more similar to that of scientific workflow, 
            particularly in a multiprocessing environment. It focused on the 
            efficient 
            scheduling of tasks and relies on I/O and network transfers to
            communicate the data between nodes. With BigData workflows, such
            operations could incur significant performance penalties. 
            To mitigate this issue and remain efficient in the face of
            BigData, Spark and Dask implement data locality, lazy evaluation 
            and in-memory processing. Data locality is the scheduler's ability
            to schedule tasks nearest to data location to reduce data transfers.
            Lazy evaluation is the workflow's ability to optimize performance
            based on the pipeline's graph structure. In-memory processing, on 
            the other hand, tries to remove any I/O that occurs between tasks. 
            Platforms such as MapReduce rely heavily 
            on I/O, but the 
            I/O performed occurs on the compute node rather than on 
            shared storaged, making it generally more efficient than scientific
            workflows. Additionally, MapReduce uses the concept of 
            locality to ensure limit data transfers across the network.
            For the efficient processing of neuroimaging BigData, it will be
            necessary for scientific worflow engines to incorporate these 
            strategies.

@@

                    As subsets of the lineage can be reused amongst different
                    structures in the pipeline, it is possible to cache the 
                    data in order to avoid any form of recomputation, which 
                    could otherwise incur performance penalties.

                \subsubsection{Data locality}
                    Data locality, with respect to BigData frameworks, is a
                    scheduling
                    strategy in which location of the data is known by the 
                    scheduler and tasks are scheduled to the location where the
                    necessary data is located. It was one of the first main 
                    strategies implemented to make BigData processing more 
                    efficient. The reasoning behind data locality is that 
                    network transfers are expensive, particularly within 
                    Big Data infrastracture, and therefore transferring large 
                    volumes of data across the network for every task would be 
                    rather expensive.

@@

                    In contrast to MapReduce, Spark and Dask do not require 
                    data to 
                    be stored on a HDFS-like filesystem. They can read data 
                    from
                    any form of shared-filesystem, including Lustre. If data 
                    is extracted from a non HDFS-like filesystem, the initial 
                    loading of data will not use data locality. However, in all
                    future tasks in which the data is already loaded into an 
                    RDD, the scheduler will be able to assign task to the most
                    favorable partition dictated by delay scheduling according
                    to data locality.

@@

                    Lazy evaluation is a strategy used throughout computer 
                    science to improve performance of operations. Rather than
                    evaluating a expression immediately, it will wait until it 
                    is required in order to compute the result. This is 
                    achieved in Spark through the calling of an \textit{action}
                    on an RDD and in Dask through the calling of a 
                    \textit{compute} function on one of its data structures.
                    By using lazy evaluation, the engines are capable of 
                    constructing graphs that result in the most optimal
                    task order of execution and further save on data transfers. 
                    \tristan{And further save on data transfers}

@@

                Alluxio~\cite{Li:2014:TRM:2670979.2670985}, 
                formerly known as Tachyon, is an in-memory filesystem
                that stores lineage data to offset the cost of writing data to 
                disk. The assumption that Alluxio makes is that it will be less
                costly to recompute the data that to store the data to disk and
                replicate across the network for fault tolerance. However,
                unlike Spark applications, Alluxio is meant to run indefinitely
                , making it unfavourable in some cases, to recompute the data
                from its starting point. For this, Alluxio evaluates the 
                lineage data and checkpoints the files which have been 
                repeatedly accessed (i.e. hot files). This strategy also helps 
                in avoiding the checkpointing of intermediary files.

                Triple-H~\cite{7152476} leverages the heterogeneity of HPC 
                clusters to 
                optimize the data's placement across the cluster for most 
                efficient use. It aims to reduce the I/O bottleneck of writing 
                through data buffering and caching. Triple-H has two modes of
                operation, Default and Lustre integrated,
                that allows it to execute on systems both with and without 
                access to a distributed filesystem. To further reduce I/O in
                HPC systems, Triple-H leverage's Lustre's fault tolerance 
                instead of using HDFS's replication-based fault-tolerance. 
                Triple-H improves 
                performance of cluster I/O by writing the data to
                the most optimal storage layer (RAMDisk, SSD, HDD, Lustre) 
                dictacted by the selected placement policy (i.e. Greedy or 
                Load-balanced). This allows alternative storage to be used as 
                a form of burst buffer to Lustre as data is written into a more
                optimal storage layer before being transferred to a slower 
                storage such as Lustre.

                Although both of these solutions have shown to improve I/O
                performance, their benefits are limited to certain 
                configurations. Alluxio needs to be installed throughout the 
                cluster occupying available memory. Data locality may also be
                and issue as the user may only be scheduled to a subset of
                nodes that do not store the data. Moreover, neuroimaging 
                operations are compute-intensive, with some tasks being 
                extremely lengthy. Recomputation may not be affordable and long
                tasks can be frequent, meaning Alluxio will need to store more 
                data that desirable.

                Triple-H's main limitation is that it assumes the compute nodes
                have local storage. This is certainly not always the case and
                any performance improvements in systems without local storage 
                may be negligable at best.

@@

            Scientific workflows typically rely on HPC schedulers for 
            distributed scheduling on clusters. These schedulers include 
            SLURM~\cite{yoo2003slurm},
            PBS~\cite{10.1007/3-540-60153-8_34}, 
            TORQUE~\cite{computing2015torque}, 
            SGE~\cite{SGE}, HTCondor~\cite{htcondor} and 
            Mesos~\cite{hindman2011mesos}. Additionally, they may provide
            their own form of scheduling for computations on a single node. 
            BigData frameworks, such as Dask,
            also provide interfaces to these HPC schedulers. However, other
            frameworks such as MapReduce and Spark are limited to Big Data 
            schedulers such as Yet Another Resource Negotiator (YARN) and MESOS.
            The reason MapReduce and Spark do not leverage HPC resource managers
            is that HPC resources managers are unaware of the locality of the 
            data and cannot perform any form of lazy evaluation.
            In addition to YARN and MEsos, Spark provides its own standalone 
            scheduler. The standalone scheduler is simple providing only a 
            basic scheduler. It allows users to use Spark out-of-the-box. Similarly
            to scientific workflows, all three BigData frameworks provide their
            own schedulers for single-node scheduling.

@@

            Since Spark does not natively interface with HPC schedulers, 
            the use of overlay Spark clusters, similar to Hadoop on Demand 
            (HoD), has been 
            investigated for Spark applications running on HPC. The reasons 
            listed for HoD's failure include poor 
            data locality when deploying a dynamic cluster on a shared HDFS 
            instance, inability to resize cluster resulting in idle nodes and 
            high cluster allocation latency. To address the issue of high 
            cluster allocation, pilot-agent scheduling strategies can be used.
            Pilot-agent scheduling strategies allow for clusters to grow 
            dynamically and to add nodes when required. This relieves users, 
            in the case of Spark, from having to wait until all nodes become 
            available to begin execution. 
            Papers~\cite{7530058}~\cite{Baer:2015:IAS:2792745.2792779} both use 
            this strategy. Furthermore, pilot-agent scheduling could be adapted
            to enable smart cluster resizing. Conversly, maintaining data 
            locality using a shared HDFS instance will remain an issue. 
            \tristan{The section is interesting, but it lacks a bit of context: one would like to 
            understand why Spark and Hadoop developed their own scheduler instead of
            reusing the ones available on HPC. And what are the main characteristics of these
            schedulers: data locality should be mentioned here.}

@@

        \section{Evaluation of BigData Frameworks for Scientific Computing}\label{eval}     
            The BigData frameworks have sparked significant interest within the 
            scientific community, leading to research on the efficiency of      
            these frameworks for scientific computing. ~\cite{souza2017spark} found  
            that Spark scales better than expected with longer tasks. This is   
            a consequence of there being less I/O occurring and the scheduler 
            not being as overloaded. 
            Additionally, ~\cite{boubela2016big} found that Spark in conjunction with  
            GPU acceleration greatly reduced the processing time of large       
            neuroimaging datasets.                                              
            As scientific, specifically neuroimaging workflows are typically    
            comprised of long tasks, it                                         
            is crucial that BigData frameworks scale well with tasks of longer  
            durations.

            To determine how different BigData frameworks compare in their 
            efficiency for the processing of scientific workflows, 
            ~\cite{mehta2017comparative} compared Spark, Dask and other Big Data 
            frameworks on
            a neuroimaging and astronomy use case. It was found that Dask, 
            and Spark performed consistently better than other frameworks and 
            had comparable 
            performance. Dask had poorer performance when data did not fit in
            memory. It was also found that Spark exhibited slower performance 
            when the individual images are larger due to the amount of transfers
            occuring between map and shuffle phases. Overall it was
            found that BigData Frameworks still required some adaptations to be
            effectively used as scientific workflow engines.

    \chapter{Portability}\label{portability}
        Platform-independence of both 
        the frameworks and tools, as well as, simple installation of all 
        necessary components and straightforward execution of workflows are a 
        requirement for frameworks and their respective pipelines to be 
        portable. Without portability, users are limited to executing certain
        frameworks only or dealing with the significant overhead of making 
        these engines work on the system available to them. In this chapter, 
        we evaluate the portability of workflows on the basis of the following
        three components: 1) platform-independence, 2) tool installation and
        3) execution setup. We also examine containerization, one of the 
        strategies used to improve portability.

        \section{Platform-independence}
            Achieving platform-independence allows users to use the workflow 
            engine in any available environment. This however does not imply 
            that the tools, and as a consequence, the pipelines created will
            also be platform-independent. Many workflow engines have opted to 
            be platform independent. These workflow engines include: Nipype,
            PSOM, SPM, FASTR, LONI, AA, Pydpiper, Taverna, MapReduce, Spark and
            Dask. These workflow engines achieve platform independence through
            running on the Java Virtual Machine (JVM) or being written in 
            Python. Platform-independent engines that run on the JVM include: 
            MATLAB engine frameworks, such as PSOM, SPM and AA; LONI; Taverna; 
            MapReduce; and Spark \tristan{this list contains Matlab
            engines}. Nipype, FastR, Pydpiper are all python-based platform 
            independent workflow engines. The remainder of the 
            workflow engines (Pegasus, Make, OpenMOLE, Nextflow, Galaxy, and 
            Toil are not platform-independent, but can execute on macOS and Linux.

            Engines built on top of a platform-independent language do not 
            necessarily imply platform independence \tristan{Grammar? Engines ... does not}. For instance, OpenMOLE is 
            built in Scala and Netflow runs on Groovy, yet neither are platform
            independent. In the case of OpenMOLE, this is due to its CARETask
            blackbox, which is only possible for Linux environments. However,
            many tools and programs available for incorporation into workflows
            are platform-dependent. For example, even though Nipype is 
            platform-independent, common tools used in Nipype modules, such as 
            FreeSurfer and FSL, are Linux-based. Moreover, HPC infrastructure 
            is usually Linux-based and it is therefore crucial that all 
            workflows are compatible with and tested on Linux systems.
